---
title: "MimeticWGCNA"
author: "Adam Stuckert"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
# load packages required for analyses
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tximport)
library(DESeq2)
library(tidyverse)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library(BiocParallel)
register(SnowParam(8)) # set the number of threads to something reasonable
library(apeglm)
library(topGO)
library(GO.db)
library(WGCNA)
library(flashClust)
library(nlme)
dir.create("results/WGCNAresults")
library(igraph)

### Note: cytoscape must be downloaded on to your local machine externally from R.
```

Print R environment information:

```{R session info}
sessionInfo()
```

This Rmarkdown file is for WGCNA analyses of the Mullerian mimicry RNA seq dataset.

#### WGCNA analyses

```{R data import}
# countdata
WGCNAdata <- read.table("data/gene.level.count.data.tsv", header = TRUE, sep = "\t")
# sample information
WGCNAsamples <- read.csv("samplespreadsheet.csv")
```

Create DESeq2 object.

```{r DESEQ creation}
# create DESeq dataset
dds <- DESeqDataSetFromMatrix(countData = WGCNAdata,
                                       colData = WGCNAsamples,
                                       design = ~ species + morph + age_weeks)

```

```{r, evaluate genes to remove}

## using counts imported via tximports from above
vsd <- vst(dds, blind = FALSE)
vsd <- assay(vsd)
# transpose, as required by WGCNA
vsd <- as.data.frame(t(vsd))

# check if there are gene outliers
gsg = goodSamplesGenes(vsd, verbose = 3)
gsg$allOK

# remove any offenders  
#if (!gsg$allOK)
#  {if (sum(!gsg$goodGenes)>0)
       #printFlush(paste("Removing genes:", paste(names(vsd)[!gsg$goodGenes], collapse= ", ")));
       #if (sum(!gsg$goodSamples)>0)
           #printFlush(paste("Removing samples:", paste(rownames(vsd)[!gsg$goodSamples], collapse=", ")))
       #vsd <- vsd[gsg$goodSamples, gsg$goodGenes]
       #}

```

Do they match up?

```{r}
# verify data and sample information align
table(WGCNAsamples$sample==rownames(vsd))

# drop lane
WGCNAsamples <- dplyr::select(WGCNAsamples, -c(filename,lane))
# make sample id row name...
WGCNAsamples <- data.frame(WGCNAsamples[,-1], row.names=WGCNAsamples[,1])

```

Cluster samples by expression.


```{R WGCNA: cluster by expression}
A = adjacency(t(vsd),type="signed") # this calculates the whole network connectivity
k = as.numeric(apply(A,2,sum))-1 # standardized connectivity
Z.k = scale(k)
thresholdZ.k = -2.5 # often -2.5
outlierColor = ifelse(Z.k<thresholdZ.k,"red","black")
sampleTree = flashClust(as.dist(1-A), method = "average")
# Convert traits to a color representation where red indicates high values
# needs to be numeric
WGCNAsamples$species <- as.numeric(WGCNAsamples$species)
WGCNAsamples$morph <- as.numeric(WGCNAsamples$morph)
WGCNAsamples$age_weeks <- as.numeric(WGCNAsamples$age_weeks)
traitColors = data.frame(numbers2colors(WGCNAsamples,signed=FALSE))
dimnames(traitColors)[[2]] = paste(names(WGCNAsamples))
datColors = data.frame(outlier = outlierColor,traitColors)

plotDendroAndColors(sampleTree,groupLabels=names(datColors),
                    colors=datColors,main="Sample Dendrogram and Trait Heatmap")

```

A few possible outliers, but we will leave it be for now.

```{R WGCNA: soft threshold power...}
powers = c(c(1:10), seq(from =10, to=30, by=1)) #choosing a set of soft-thresholding powers
sft = pickSoftThreshold(vsd, powerVector=powers, verbose =5, networkType="signed") #call network topology analysis function

```

```{R WGCNA: scale independence}
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab= "Soft Threshold (power)", ylab="Scale Free Topology Model Fit, signed R^2", type= "n", main= paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers, col="red")
abline(h=0.90, col="red")
```

```{R WGCNA: connectivity}
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab= "Soft Threshold (power)", ylab="Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")
```

Soft power asymptotes pretty hard. I'll attempt this with 20. ### Note to self is this right????

Construct networks.

```{R WGCNA: construct networks}
net = blockwiseModules(vsd, power = 10,
    TOMType = "unsigned", minModuleSize = 30,
    reassignThreshold = 0, mergeCutHeight = 0.25,
    numericLabels = TRUE, pamRespectsDendro = FALSE,
    saveTOMs = TRUE,
    saveTOMFileBase = "MimeticGenesTOM",
    verbose = 3)

```

Plot these modules.

```{R WGCNA: plot modules}
table(net$colors)

# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
    "Module colors",
    dendroLabels = FALSE, hang = 0.03,
    addGuide = TRUE, guideHang = 0.05)
```



Correlate traits.

```{R WGCNA: correlate traits}
#Define number of genes and samples
nGenes = ncol(vsd)
nSamples = nrow(vsd)

#Recalculate MEs with color labels
MEs0 = moduleEigengenes(vsd, mergedColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, WGCNAsamples, use= "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)

#Print correlation heatmap between modules and traits
textMatrix= paste(signif(moduleTraitCor, 2), "\n(", 
                    signif(moduleTraitPvalue, 1), ")", sep= "")
dim(textMatrix)= dim(moduleTraitCor)
par(mar= c(6, 8.5, 3, 3))
#display the corelation values with a heatmap plot
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= names(WGCNAsamples), 
            yLabels= gsub(pattern = "ME", replacement = "", names(MEs)), 
            ySymbols= names(MEs), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(50), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 0.5, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))


# save to disk

png("figures/WGCNA_heatmap.png", width = 9.5, height = 7, units = "in", res = 300)
par(mar= c(6, 12, 3, 3))
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= names(WGCNAsamples), 
            yLabels= gsub(pattern = "ME", replacement = "", names(MEs)), 
            ySymbols= gsub(pattern = "ME", replacement = "", names(MEs)), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(50), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 0.5, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))
dev.off()
```



Get module membership.


```{r WGCNA: output module membership}

datME=moduleEigengenes(vsd,mergedColors)$eigengenes
datKME=signedKME(vsd, datME, outputColumnName="MM.")
genes=names(vsd)
geneInfo0 = data.frame(gene=genes,moduleColor=mergedColors)
color=data.frame(geneInfo0,datKME) #these are from your original WGCNA analysis 
head(color)
write.csv(as.data.frame(color), file = "results/GeneModule_membership.csv")

#### MM pvalues
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(vsd, MEs, use = "p"));
MMPvalue=as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
pvals=data.frame(geneModuleMembership,MMPvalue)
head(pvals)
write.csv(as.data.frame(pvals), file = "results/GeneModule_membership_pvalues.csv")

```

Gene significance by module membership.


```{R WGCNA: significance of population membership}

whichTrait="morph" #Replace this with the trait of interest

morph = as.data.frame(WGCNAsamples$morph);
names(morph) = "morph"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(vsd, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(vsd, morph, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(morph), sep="");
names(GSPvalue) = paste("p.GS.", names(morph), sep="");

```



PULL OUT datKME by module of interest, rank them high membership to lowest (from 1 to -1), then run through GO.

For each signifcant module I want to see:
1. What genes are in these modules.
2. If there are candidate color genes.
3. GO of this grouping.

### Significant gene modules by color morph

```{R, WGNCA: identify significant modules by morph}
# subets out only significant modules
modulePs <- as.data.frame(moduleTraitPvalue)
morphsigmods <- rownames(subset(modulePs, morph < 0.05))
print(paste0("Number of significant modules by color morph: ", length(morphsigmods)))
speciessigmods <- rownames(subset(modulePs, species < 0.05))
print(paste0("Number of significant modules by species: ", length(speciessigmods)))
stagesigmods <- rownames(subset(modulePs, age_weeks < 0.05))
print(paste0("Number of significant modules by developmental stage: ", length(stagesigmods)))

sigmods <- rownames(modulePs[rowSums(modulePs < 0.05) >= 1, ])
print(paste0("Number of unique significant modules by any variable: ", length(sigmods)))

# get info for any of these with p values (example = stages):
# format(subset(modulePs, morph < 0.05), scientific = FALSE)
# 
```

## What are my hub genes?

```{R Hub genes}

# remove "ME" from all sigmods
sigmods <- gsub(pattern = "ME", replacement = "", sigmods)
probes = names(vsd)
nTop = 30
moduleColors=mergedColors
adj = adjacency(vsd, power = 10, type = "signed") #specify network type

## Translate adjacency into topological overlap matrix and calculate the corresponding dissimilarity

TOM = TOMsimilarity(adj, TOMType="signed") #Topological Overlap Matrix
dissTOM = 1-TOM 

for(i in 1:length(unique(sigmods))){
  #inModule = sigmods[i];
  #modProbes = probes[inModule];
  inModule = (moduleColors == sigmods[i]);
  modProbes = probes[inModule];
  ## Select the corresponding Topological Overlap
  
  modTOM = TOM[inModule, inModule];
  dimnames(modTOM) = list(modProbes, modProbes)
  
  ## Genes with highest connectivity within each significant module
  
  IMConn = softConnectivity(vsd[, modProbes])
  top = (rank(-IMConn) <= nTop)
  
  tempdat<-data.frame(IMConn,modProbes)[order(-IMConn),]
  names(tempdat)<-c("IMConnectivity","target_id")
  tempdat$ModuleColor <- rep(sigmods[i],nrow(tempdat))
  #tempdat2 = left_join(tempdat,preGO1, by = "target_id")
  write.csv(tempdat, paste("results/WGCNAresults/HubGenes",sigmods[i],".csv",sep="_"), na="", row.names=FALSE)
}

## List all the files

temp <- list.files(path = "results/WGCNAresults", pattern = "*HubGenes")
temp <- paste("results/WGCNAresults/", temp, sep = "")
Hub_genes = do.call(rbind, lapply(temp, function(x) read.csv(x, stringsAsFactors = FALSE)))
write.csv(Hub_genes, paste("results/WGCNAresults/HubGenes_all.csv",sep="_"), na="", row.names=FALSE)

```

```{R HUBGENES2}

# Recalculate topological overlap if needed
#TOM = TOMsimilarityFromExpr(vsd, power = 10)

# Select modules
sigmods # Lists modules of interest
module = "purple"    
moduleColors = mergedColors
# Select module probes
probes = names(vsd)
inModule = is.finite(match(moduleColors, module));
modProbes = probes[inModule];
#modGenes = annote_df$gene_name[match(modProbes, annote_df$target_id)];

# Select the corresponding Topological Overlap
modTOM = TOM[inModule, inModule];
dimnames(modTOM) = list(modProbes, modProbes)
cyt = exportNetworkToCytoscape(modTOM,
       edgeFile = paste("results/WGCNAresults/CytoscapeInput-edges-", paste(module, collapse="-"), ".txt"),
       nodeFile = paste("results/WGCNAresults/CytoscapeInput-nodes-", paste(module, collapse="-"), ".txt"),
       weighted = TRUE,
       threshold = 0.175, # Change to get about 30
       nodeNames = modProbes,
       #altNodeNames = modGenes,
       nodeAttr = moduleColors[inModule]);
cyt

## Record of thresholds:
# Dark grey - 0.02
# light green - 0.1
# blue - 0.4
# darkgreen - 0.02
# darkturquoise - 0.1
# red = 0.4
# turquoise = 0.4
# salmon = 0.15


cytoscapePing()

```


### Rank genes from significant modules for GO analyses
Get gene information for these.

```{R, WGNCA: gene information for significant modules by morph}
MMmorphsigmods <- gsub(pattern = "ME", replacement = "MM.", morphsigmods)
#setDF(datKME)
#datKME %>% dplyr::select(datKME, matches(paste(MMsigmods, collapse="|")))
#datKME[,c(sigmods)]

#sigmorphmodules <- datKME[,c("MM.pink", "MM.purple", "MM.grey")]


# loop to make ordered outputs for GOrilla
for (i in 1:length(MMmorphsigmods)){
  modcol <- MMmorphsigmods[i]
  # get data for module color
  tmp <- datKME[,modcol]
  gene_rows <- rownames(datKME)
  tmp <- as.data.frame(cbind(gene_rows, as.numeric(tmp)))
  colnames(tmp) <- c("gene", "module.membership")
  tmp$module.membership <- as.numeric(as.character(tmp$module.membership))
  tmp <- tmp[order(tmp[2], decreasing = TRUE), ]
  write.csv(tmp, paste0("results/WGCNAresults/morphs_", modcol, "ModuleMembership.csv"), row.names = FALSE)
}
```

```{R, WGNCA: gene information for significant modules by morph}
MMspeciessigmods <- gsub(pattern = "ME", replacement = "MM.", speciessigmods)
#setDF(datKME)
#datKME %>% dplyr::select(datKME, matches(paste(MMsigmods, collapse="|")))
#datKME[,c(sigmods)]

#sigmorphmodules <- datKME[,c("MM.pink", "MM.purple", "MM.grey")]


# loop to make ordered outputs for GOrilla
for (i in 1:length(MMspeciessigmods)){
  modcol <- MMspeciessigmods[i]
  # get data for module color
  tmp <- datKME[,modcol]
  gene_rows <- rownames(datKME)
  tmp <- as.data.frame(cbind(gene_rows, as.numeric(tmp)))
  colnames(tmp) <- c("gene", "module.membership")
  tmp$module.membership <- as.numeric(as.character(tmp$module.membership))
  tmp <- tmp[order(tmp[2], decreasing = TRUE), ]
  write.csv(tmp, paste0("results/WGCNAresults/species_", modcol, "ModuleMembership.csv"), row.names = FALSE)
}
```

```{R, WGNCA: gene information for significant modules by morph}
MMstagesigmods <- gsub(pattern = "ME", replacement = "MM.", stagesigmods)
#setDF(datKME)
#datKME %>% dplyr::select(datKME, matches(paste(MMsigmods, collapse="|")))
#datKME[,c(sigmods)]

#sigmorphmodules <- datKME[,c("MM.pink", "MM.purple", "MM.grey")]


# loop to make ordered outputs for GOrilla
for (i in 1:length(MMstagesigmods)){
  modcol <- MMstagesigmods[i]
  # get data for module color
  tmp <- datKME[,modcol]
  gene_rows <- rownames(datKME)
  tmp <- as.data.frame(cbind(gene_rows, as.numeric(tmp)))
  colnames(tmp) <- c("gene", "module.membership")
  tmp$module.membership <- as.numeric(as.character(tmp$module.membership))
  tmp <- tmp[order(tmp[2], decreasing = TRUE), ]
  write.csv(tmp, paste0("results/WGCNAresults/developmentalstage_", modcol, "ModuleMembership.csv"), row.names = FALSE)
}
```


